---
title: "Final Report"
author: "Mark JP Sanchez"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = FALSE,
  include = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

# Introduction

League of Legends is a hugely popular video game that rakes in billions of dollars a year and is easily one of the most popular and influential games in the world. The success of the game has even spawned a successful e-sports scene, a k-pop band, and even a critically acclaimed Netflix show. All of this, spawning from an online 5 vs. 5 competitive MOBA game. A MOBA game effectively being a game such that two teams face off head-to-head in a match. Each player chooses a character, in League of Legends they are called Champions, to play as and each character has a unique set of skills. During the game, each player attempts to level up their character, accrue money to spend on items, destroy enemy towers and escort their minions to the enemy base. The objective of the game is to destroy your opponent's base but to do so, you need to power up and work with your team to both infiltrate their base and protect yours over the course of 20 minutes to an hour. 

Evidently. one thing that makes the game so much fun is all the moving parts and decisions a player can make to influence a match. Some decisions include what champion to choose, what items to buy, and what objective to focus on. Unfortunately, all of these factors and decisions can overwhelm new and existing players, and with the main objective obviously being to win the game, it would be nice to now what kind of decisions actually influence those wins.

A link to the full report can be found [here](report/jsc370_final.pdf)

# Methods

## Data Wrangling and Cleaning Methods

The data set was taken from the RIOT api and contains data from League of Legends matches. The data collected contains info such as how long the matches were, what champions each player played, how much damage each player did, and what side objectives the players took. The data was given in JSON format which I had to turn into nice R data frames.

One issue that arose from gathering match data from the RIOT api was that there was no proper endpoint for gathering random matches. The main way to gather matches was to call an endpoint that gave you a list of past matches of a specific player. The problem with this is that if I collected all of my match data from this list of matches, then I would have a confounding variable being this one player. I want to make my data between matches as independent as possible. Luckily, each match consists of 10 players so I can recursively gather matches of each player in the current match. Using this method, I gathered the my match data by first gathering a few thousand player ids by recursively checking the players in a given player's latest match. I then get the latest match before a previous date of each player. My chosen date was Valentines day which was a week before I gathered my data. There was also a large update to the game on January 7th, 2022 so we need to make sure our matches don't come before then. This should give me a list of matches that has little overlap in players.

Calling the match data api from RIOT gives you a lot of variables to work with. Especially since each player in the match have tens of different variable attached to them. With 10 players per game, that is a lot of data to work with. Since we are looking at the decisions each player makes on their own, we will take the data and split it up into 2 different datasets. The first dataset will contain match metadata such as when the match was created and how long the match was. The second dataset contains data on each player during a match. So each observation in that player dataset will be corresponding to data from a player in a match. This data will include things like how much damage they did, how many kills they got and how much gold they made.

With the methodologies I listed above I ran my code and let the program run for a few hours. This long waiting time was due to the rate limit imposed by the RIOT api of a max of 100 requests per 2 minutes. In the end we collected a dataset of 9400 matches.

```{r load-data}
# Loading in our data
player_df <- read.csv("data/lol_player_info.csv")
match_df <- read.csv("data/lol_match_info.csv")

# Getting rid of games that are too old
match_df <- match_df %>%
  filter(gameCreation > 1641621600000)

match_df$gameDate <- as.Date(as.POSIXct(match_df$gameCreation / 1000, origin="1970-01-01"))
match_df$gameDuration <- match_df$gameDuration / 60

player_df <- player_df %>%
  mutate(
    teamPosition = ifelse(teamPosition == "", NA, teamPosition)) %>%
  drop_na(teamPosition)

player_df <- match_df %>%
  select(match_id) %>%
  merge(player_df, by ="match_id", all.x = T)

avged_stats <- player_df %>%
  merge(match_df, by = "match_id") %>%
  mutate(kills = kills/ gameDuration,
         totalDamageDealt = totalDamageDealt / gameDuration,
         totalHeal = totalHeal / gameDuration,
         wardsPlaced = wardsPlaced / gameDuration,
         totalTimeCCDealt = totalTimeCCDealt / gameDuration,
         totalDamageTaken = totalDamageTaken / gameDuration,
         goldEarned = goldEarned / gameDuration,
         win_state = ifelse(win, "Won", "Lost"))
```

# Interactive Visualization

To get a better sense of the data here are some interactive visuals of the dataset

## Champion Info

One of the most notable and recognizable things about League of Legends is there character, also known as Champions. Here we have some graphs that give us some insight on some basic stats on each Champion.

```{r plot1, fig.height=10, include = T}

player_df %>%
  group_by(championName) %>%
  summarise(play_rate = n()/nrow(match_df) * 100) %>%
  plot_ly(x = ~championName,
          y = ~play_rate,
          type = "bar",
          color = ~championName,
          hover_info = "text",
          text = ~paste(paste0("Champion Name: ", championName),
                        paste0("Play Rate: ", paste0(round(play_rate,2), "%")),
                        sep = "<br>"
                        )) %>%
  layout(title = "Play Rate of Each Champion (%)",
         yaxis = list(title = "Play Rate (%)"),
         xaxis = list(title = "Champion"))
```

Here play rate is defined as the percentage of games in our data set that the champion was present in. Notice that since there are 10 champions per team, the play rate won't add up to 100%

```{r plot2, fig.height=10, include = T}

player_df %>%
  group_by(championName) %>%
  summarise(win_rate = sum(win)/n() * 100, n = n()) %>%
  plot_ly(x = ~championName,
          y = ~win_rate,
          type = "bar",
          color = ~championName,
          hover_info = "text",
          text = ~paste(paste0("Champion Name: ", championName),
                        paste0("Win Rate: ", paste0(round(win_rate,2), "%")),
                        paste0("Out of ", paste0(n, " games")),
                        sep = "<br>"
                        )) %>%
  layout(title = "Win Rate of Each Champion (%)",
         yaxis = list(title = "Win Rate (%)"),
         xaxis = list(title = "Champion"))
```

## Match and Player Stats Info

Here are some graphs and info on matches and different player stats

```{r plot7, fig.height=10, include = T}
diff_stats <- match_df %>%
  left_join(avged_stats, by = "match_id") %>%
  mutate(win = win_state) %>%
  group_by(match_id, teamId, win) %>%
  summarise(
    damageDealt = sum(totalDamageDealt),
    goldEarned = sum(goldEarned)
    ) %>%
  pivot_wider(names_from = teamId, values_from = c("damageDealt", "goldEarned", "win")) %>%
  mutate(
    damageDealtDiff = damageDealt_100 - damageDealt_200,
    goldEarnedDiff = goldEarned_100 - goldEarned_200
    )

diff_stats %>%
  plot_ly(
    x = ~damageDealtDiff,
    color = ~win_100,
    type = "histogram"
  ) %>%
  layout(title = "Histogram of Differences in Damage Dealt Per Minute by Win/Loss",
         yaxis = list(title = "Frequency"),
         xaxis = list(title = "Damage Dealt Difference Per Minute"),
         barmode = "stack")
```

```{r plot8, fig.height=10, include = T}
diff_stats %>%
  plot_ly(
    x = ~goldEarnedDiff,
    color = ~win_100,
    type = "histogram"
  ) %>%
  layout(title = "Histogram of Differences in Gold Earned Per Minute by Win/Loss",
         yaxis = list(title = "Frequency"),
         xaxis = list(title = "Gold Earned Difference Per Minute"),
         barmode = "stack")
```

## Latent Factors

Within our report we used some matrix factorization to learn some latent factors of each of the 158 champions. Our matrix was 158x158 where each row and column represented a champion. With our xgboost model we learned that gold earned is one of the biggest indicators of who is winning the match. So for our matrix, each cell represents the average amount of gold earned per minute in games where the column champion is on the opposing team. Here we can gather information on an important metric as well and capture any interactions between the champion and its opponents.

Our model gave us an r squared value of 0.88 so our model explains 88% of the variance in our test set. Below is an example of the champions by their latent factors.

```{r}
champion_matrix_df <- read.csv("data/champion_matrix.csv")

champion_matrix_df %>%
  plot_ly(x = ~attacking_factor1,
          y = ~attacking_factor2,
          type = "scatter",
          mode = "markers",
          color = ~teamPosition,
          hover_info = "text",
          text = ~paste("Champion Name: ", champion
                        )) %>%
  layout(title = "Learned Latent Factors of Champions",
         yaxis = list(title = "Latent Factor 2"),
         xaxis = list(title = "Latent Factor 1"))
```

# Conclusion and Summary

From fitting our model and looking at the data, we can gather a lot about the game of League of Legends. Our initial exploratory graphs gave us insight on the relationship of variables, our xgboost model showed us what skills are important and what aren't, and our matrix factorization gave us an understanding of how to make your choice on what champion to choose.

There are many decisions to be made and many skills to polish which can be overwhelming for a new player. So to aid new and veteran players, here is our recommendation on how you can win a game in League of Legends,

- You should polish your skills in earning as much gold as possible. This means killing and last-hitting minions, killing enemies and reducing any downtime you have. Use gold, and the difference of gold between you and your opponents, as a benchmark on who is currently doing better in the game
- Focus on getting kills and staying alive. This seems pretty evident but there are some nuances. Focus on getting high killing sprees. This means that just because you can get a kill doesn't mean you should. How likely is death if you attempt to secure a kill? If you can guarantee a kill at the expense of your own life, you should reconsider your options. That doesn't mean never trade your lives for another player but if you already have a number of kills under your belt, be wary of dangerous situations.
- Kills are more important than doing damage. While doing damage is useful, you are far more likely to win a game the more kills you have. So even if it feels like overkill, unless you leave yourself defenseless, do not be ashamed of unloading more than enough abilities on an enemy to ensure they die.
- Do not feel too pressured to choose the "best" or most "meta" champion to play. All of the champions seem to be fairly balanced and more notably distinct from each other. Choose a role you want to play and a champion that fits that role. Since there is some distinction between each champion it is better to choose specific champion and learn to get good with them. While some skills are transferable between champions, each one of them has their own flavour.If you are new and attempt to learn whatever is currently the "best" current champion, then you might find that your progress will be slowed down due to you constantly switching between differing characters.
- While being a team game, your personal achievements and actions within the game are very influential in the outcome. Even if you feel like your team is horrible, you are not guaranteed to lose. If you try your best, you might find that victory is possible.

Our last recommendation is one that needs no machine learning model support it. The last recommendation being have fun! At the end of the day, League of Legends is a video game and the goal of a video game is enjoyment. While it is cool to win, it is impossible to win every game so just have fun as you improve your skills.
