---
title: "DataSet Preview"
author: "Mark JP Sanchez"
output: 
  html_document:
    toc: True
    toc_float: True
---

<br>

This is my JSC370 Final Project website. I will showcase a few interactive visuals here.

<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = FALSE,
  include = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```


```{r load-data}
# Loading in our data
player_df <- read.csv("data/lol_player_info.csv")
match_df <- read.csv("data/lol_match_info.csv")

# Getting rid of games that are too old
match_df <- match_df %>%
  filter(gameCreation > 1641621600000)

match_df$gameDate <- as.Date(as.POSIXct(match_df$gameCreation / 1000, origin="1970-01-01"))
match_df$gameDuration <- match_df$gameDuration / 60

player_df <- player_df %>%
  mutate(
    teamPosition = ifelse(teamPosition == "", NA, teamPosition)) %>%
  drop_na(teamPosition)

player_df <- match_df %>%
  select(match_id) %>%
  merge(player_df, by ="match_id", all.x = T)

avged_stats <- player_df %>%
  merge(match_df, by = "match_id") %>%
  mutate(kills = kills/ gameDuration,
         totalDamageDealt = totalDamageDealt / gameDuration,
         totalHeal = totalHeal / gameDuration,
         wardsPlaced = wardsPlaced / gameDuration,
         totalTimeCCDealt = totalTimeCCDealt / gameDuration,
         totalDamageTaken = totalDamageTaken / gameDuration,
         longestTimeSpentLiving = longestTimeSpentLiving / gameDuration,
         goldEarned = goldEarned / gameDuration,
         win_state = ifelse(win, "Won", "Lost"))
```

# Interactive Visualization

To get a better sense of the data here are some interactive visuals of the dataset, (bonus graphs)

## Match and Player Stats Info

Here are some graphs and info on matches and different player stats

```{r plot6, fig.height=10, include = T}
match_df %>%
  plot_ly(
    x = ~gameDuration,
    type = "histogram"
  ) %>%
  layout(title = "Histogram of Game Lengths",
         yaxis = list(title = "Frequency"),
         xaxis = list(title = "Game Length in Minutes"))
```

```{r plot7, fig.height=10, include = T}
avged_stats %>%
  filter(longestTimeSpentLiving > 0) %>%
  mutate(longestTimeSpentLiving = longestTimeSpentLiving / 60 / 60) %>%
  plot_ly(
    x = ~longestTimeSpentLiving,
    type = "histogram"
  ) %>%
  layout(title = "Histogram of Longest Time Spent Living (%)",
         yaxis = list(title = "Frequency"),
         xaxis = list(title = "Longest Time Player Was alive as % of game length"))
```

```{r plot8, fig.height=10, include = T}
avged_stats %>%
  plot_ly(
    x = ~largestKillingSpree,
    type = "histogram"
  ) %>%
  layout(title = "Histogram of Largest Killing Spree",
         yaxis = list(title = "Frequency"),
         xaxis = list(title = "Largest Killing Spree"))
```

## Role Info

There are 5 distinct roles each player plays in a game. The 5 positions are Bottom, Jungle, Middle, Top and Utility. Here we will show some graphs to give a general idea on what each position does.

```{r plot3, fig.height=10, include = T}
avged_stats <- player_df %>%
  merge(match_df, by = "match_id") %>%
  mutate(kills = kills/ gameDuration,
         totalDamageDealt = totalDamageDealt / gameDuration,
         totalHeal = totalHeal / gameDuration,
         wardsPlaced = wardsPlaced / gameDuration,
         totalTimeCCDealt = totalTimeCCDealt / gameDuration,
         totalDamageTaken = totalDamageTaken / gameDuration,
         goldEarned = goldEarned / gameDuration,
         win = ifelse(win, "Won", "Lossed"))

avged_stats %>%
  drop_na(teamPosition) %>%
  plot_ly(
    y = ~totalDamageDealt,
    color = ~teamPosition,
    type = "box"
  ) %>%
  layout(title = "Damage Dealt Per Minute By Role",
         yaxis = list(title = "Damage Dealt Per Minute"),
         xaxis = list(title = "Role"))
```

```{r plot4, fig.height=10, include = T}
avged_stats %>%
  drop_na(teamPosition) %>%
  plot_ly(
    y = ~totalDamageTaken,
    color = ~teamPosition,
    type = "box"
  ) %>%
  layout(title = "Damage Taken Per Minute By Role",
         yaxis = list(title = "Damage Taken Per Minute"),
         xaxis = list(title = "Role"))
```

```{r plot5, fig.height=10, include = T}
avged_stats %>%
  drop_na(teamPosition) %>%
  plot_ly(
    y = ~goldEarned,
    color = ~teamPosition,
    type = "box"
  ) %>%
  layout(title = "Gold Earned Per Minute By Role",
         yaxis = list(title = "Gold Earned Per Minute"),
         xaxis = list(title = "Role"))
```

